### 2023-01-16 paper.tex
add latex makefile and paper abstract

\documentclass[12pt]{amsart}

\usepackage{hyperref}

\title{Wiggling the bandwidth in kernel ridgeless regression 
    with scale-dependent kernels}
\author{Luca Wellmeier}
\begin{document}
\begin{abstract}
    We explore the possibility of using two different bandwidths in the 
    fitting and evaluation part in kernel ridgeless regression.
    The provided experimental results indicate that \textit{wiggling}
    the bandwidth in this way can generalize and be a cheap alternative to
    the classical Tikhonov square-norm penality.
    Finally, we propose and test a new iterative method for optimizing the 
    bandwidth parameter by employing wiggling in a local search.
\end{abstract}
\maketitle
\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Wiggling as a regularization technique}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Iterative wiggling to find the best bandwidth}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and directions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2023-01-16 paper.tex
add basic latex packages
add review and background on kernel ridgeless estimators

@@
 
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
 \usepackage{hyperref}
+\usepackage[nopatch=eqnum]{microtype}
+\usepackage{cleveref}
 
-\title{Wiggling the bandwidth in kernel ridgeless regression 
+\newcommand{\R}{\mathbb{R}}
+
+\title{Wiggling the bandwidth in ridgeless regression 
     with scale-dependent kernels}
@@
 
+Let $\mathbf X \in \R^{n \times d}$ be a matrix of $n$ training examples
+$X_1, \dots, X_n \in \R^d$ in $d$-dimensional Euclidean space as rows and 
+let $Y$ be the column vector of responses $y_1, \dots, y_n \in \R$.
+The main actor of this paper is the kernel ridgeless regression estimator
+that predicts the response on an unseen data point $X \in \R^d$ as
+\begin{equation} \label{eq:ridgeless}
+    \hat f(X) = \sum_{i=1}^n (\mathbf K^\dagger Y)_i K(X, X_i).
+\end{equation}
+Here, $mathbf K$ denotes the kernel matrix $\mathbf K_{ij} = k(X_i, X_j)$
+is the \emph{kernel matrix} obtained by evaluating a kernel function 
+$k \colon \R^d \times \R^d \to \R$ on the training examples and 
+$\mathbf K^\dagger$ denotes its \emph{pseudo-inverse}.
+We briefly recall the motivations behind these notions.
+First, the \emph{representer theorem} asserts that any minimizer to the 
+least squares problem
+\[ \min_{f \in \mathcal H_k} \frac 1n \sum_{i=1}^n (y_i - f(X_i))^2 \]
+in the \emph{reproducing kernel Hilbert space} $\mathcal H_k$ associated
+to $k$, is given by a linear combination of the form
+$\sum_{i=1}^n c_i k(X,X_i)$.
+This allows to apply the usual linear least squares method to a - at the 
+first glance - non-linear approximation problem in a possibly 
+infinite-dimensional hypothesis space by pushing the data points into 
+the \emph{feature space} spanned by $K(X, \cdot) \in \mathcal H_k$ 
+for $X \in \R^d$.
+The kernel matrix in this setting corresponds to the empirical covariance
+matrix $\mathbf X^T \mathbf X$ and classical theory suggests that, even if 
+it does not possess full-rank, there will always exist a minimizer to 
+the least squares problem above.
+The pseudo-inverse then corresponds to the unique least-squares solution
+with minimal $\mathcal H_k$-norm.
+This reduction of an infinite-dimensional optimization problem to a
+finite-dimensional one using kernels is generally known as the 
+\emph{kernel trick}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2023-01-16 code/mnist.py
add mnist loader

@@ -1,0 +1,43 @@
+import os
+import numpy as np
+from sklearn.datasets import fetch_openml
+
+MNIST_PATH = 'build/mnist_784.npz'
+
+class MnistBinaryDataset:
+
+    def __init__(self, seed):
+        if not os.path.exists(MNIST_PATH):
+            mnist = fetch_openml('mnist_784', parser='auto')
+            self.allX = mnist.data.to_numpy()
+            self.allY = mnist.target.astype(int).to_numpy()
+            np.savez(MNIST_PATH, X=self.allX, Y=self.allY)
+        else:
+            archive = np.load(MNIST_PATH)
+            self.allX = archive['X']
+            self.allY = archive['Y']
+        self.dim = self.allX.shape[1]
+        self.rng = np.random.default_rng(seed)
+
+    def sample(self, digitA, digitB, split):
+        assert(np.all(np.array(split) > 0))
+        relevant = np.logical_or(self.allY == digitA, self.allY == digitB)
+        N = np.sum(split)
+        assert(np.sum(relevant) >= N)
+
+        perm = self.rng.permutation(N)
+        X = self.allX[relevant][:N][perm]
+        Y = self.allY[relevant][:N][perm].reshape(-1,1)
+        Y[Y == digitA] = -1
+        Y[Y == digitB] = 1
+
+        curPos = 0
+        r = []
+        for p in split:
+            r.append(X[curPos:curPos+p])
+            r.append(Y[curPos:curPos+p])
+            curPos += p
+        return r
+
+if __name__ == '__main__':
+    MnistBinaryDataset(0)

### 2023-01-16 paper.tex
add bibliography and write interpolating regime motivations

@@
 \section{Introduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 Let $\mathbf X \in \R^{n \times d}$ be a matrix of $n$ training examples
@@
 \end{equation}
-Here, $mathbf K$ denotes the kernel matrix $\mathbf K_{ij} = k(X_i, X_j)$
+Here, $\mathbf K$ denotes the kernel matrix $\mathbf K_{ij} = k(X_i, X_j)$
 is the \emph{kernel matrix} obtained by evaluating a kernel function 
@@
 
+Recently, there was a new wave of interest into kernels sparked by the 
+fundamental question of why highly over-parameterized deep neural nets
+perform as well as they do given that the usual established model, 
+the \emph{bias-variance trade-off}, does not capture this regime.
+The idea is that to find a good estimator one should choose the 
+complexity of the hypothesis space should be high enough to avoid 
+\emph{underfitting} but simultaneously low enough to avoid 
+\emph{overfitting}.
+For instance, if we roughly identify the number of adjustable parameters
+of a model with the complexity of the hypothesis space, we would expect
+a U-shaped test error curve over the number of parameters with very high
+risk at the extrema where $N_{\textrm{param}}$ is very small but also
+$N_{\textrm{param}} \approx N_{\textrm{train}}$, where the model is able
+to \emph{interpolate} the training data - it learns them by heart and 
+the analogy is the same as in the human experience where learning by heart
+does not give you true understanding.
+However, modern, practically successful, deep neural nets tend to have a
+parameter count that exceeds the training sample count by orders of 
+magnitude.
+This is not captured by classical theory so the question is two-fold:
+is this the reason for their performance and 
+what exactly happens in this \emph{interpolating regime}.
+A good start into the literature would be the following papers on 
+the \emph{double descent} curve (\cite{doubledescent}) and on the new 
+concept of \emph{benign overfitting} (\cite{benignoverfitting}).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Wiggling as a regularization technique}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \section{Iterative wiggling to find the best bandwidth}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \section{Conclusions and directions}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{thebibliography}{20}
 
+\bibitem{doubledescent}
+Belkin, Mikhail, et al. "Reconciling modern machine learning practice and the bias-variance trade-off." arXiv preprint arXiv:1812.11118 (2018).
+
+\bibitem{benignoverfitting}
+Bartlett, Peter L., et al. "Benign overfitting in linear regression." Proceedings of the National Academy of Sciences 117.48 (2020): 30063-30070.
+
+\bibitem{understandkernels}
+Belkin, Mikhail, Siyuan Ma, and Soumik Mandal. "To understand deep learning we need to understand kernel learning." International Conference on Machine Learning. PMLR, 2018.
+
+\bibitem{justinterpolate}
+Liang, Tengyuan, and Alexander Rakhlin. "Just interpolate: Kernel “ridgeless” regression can generalize." The Annals of Statistics 48.3 (2020): 1329-1347.
+
+\end{thebibliography}
 \end{document}

### 2023-01-16 paper.tex
motivate use of gaussian and laplacian

@@
 the \emph{bias-variance trade-off}, does not capture this regime.
-The idea is that to find a good estimator one should choose the 
-complexity of the hypothesis space should be high enough to avoid 
+The idea is that in order to find a good estimator one should choose 
+the complexity of the hypothesis space to be high enough to avoid 
 \emph{underfitting} but simultaneously low enough to avoid 
@@
 A good start into the literature would be the following papers on 
-the \emph{double descent} curve (\cite{doubledescent}) and on the new 
+the \emph{double descent} phenomenon (\cite{doubledescent}) and on the
 concept of \emph{benign overfitting} (\cite{benignoverfitting}).
+
+In particular, there are many indications that kernel learning methods 
+behave analgously to deep learning from many different perspectives.
+Much evidence was provided in \cite{understandkernels}.
+First, many kernels as for instance Laplacian and Gaussian are always 
+able to interpolate the training data perfectly, providing a prime example
+of interpolating estimators.
+Unlike deep neural nets, kernel methods are easy to handle not only on 
+paper but also practically due to the very small amount of hyper-parameters
+and there are many tasks in which their performance is comparable to the 
+latter or even exceed it.
+In \cite{justinterpolate}, the authors provide experimental and theoretic 
+evidendence that it is better to interpolate instead of the classically 
+suggested norm penalties (\textit{à la} Tikhonov) under certain circumstances.
+Now, these circumstances have not been fully understood but some necessary 
+conditions seem to have pinned down:
+high-dimensionality (\cite{laplaceconsistency,benignoverfitting}) 
+with low effective dimension (\cite{benignoverfitting})
+and properly chosen smoothness of the kernel w.r.t. the nature of the data 
+(\cite{justinterpolate}).
+Moreover, there seems to be strong links (\cite{ntklaplacian}) between 
+Laplacian kernels and ReLU feed-forward neural nets through another kernel 
+called the \emph{neural tangent kernel} that is able to capture the 
+learning dynamics of gradient descent there.
+
+These reasons have motivated our choice to consider here the extrema of 
+the spectrum of \emph{Matérn kernels}: a family of kernels whose associated
+RKHSs are exactly Sobolev spaces with real smoothness index.
+The least smooth member of the family, the \emph{Laplacian kernel}, is 
+defined by
+\[ k^L_\gamma(x,x') = \exp(-\gamma \| x - x' \|) \]
+and the smoothest member of it is the \emph{Gaussian kernel}
+\[ k^G_\gamma(x,x') = \exp(-\gamma \| x - x' \|^2). \]
 
@@
 
+\bibitem{ntklaplacian}
+Geifman, Amnon, et al. "On the similarity between the laplace and neural tangent kernels." Advances in Neural Information Processing Systems 33 (2020): 1451-1461.
+
+\bibitem{laplaceconsistency}
+Rakhlin, Alexander, and Xiyu Zhai. "Consistency of interpolation with Laplace kernels is a high-dimensional phenomenon." Conference on Learning Theory. PMLR, 2019.
+
 \end{thebibliography}

### 2023-01-17 code/exp_iterative_algo.py
add figure pipeline and add iterative algorithm experiment

@@ -1,0 +1,51 @@
+import numpy as np
+from sklearn.metrics import mean_squared_error
+from matplotlib import pyplot as plt
+from code.shared import MnistBinaryDataset, laplace_kernel, IteratedWiggleSearchScaleKernelLS
+
+RES_FILE = 'build/results/iterative_algo.npz'
+FIG_FILE = 'build/figures/iterative_algo.png'
+
+Ntr = 200
+Nte = 500
+Xtr,Ytr, Xte,Yte = MnistBinaryDataset(3281).sample(4, 9, [Ntr, Ntr])
+
+gamma0 = 1/784
+n_iter = 8
+search_n = 5
+search_loexp = 2
+split = 0.85
+
+def compute():
+    model = IteratedWiggleSearchScaleKernelLS(kernel_func=laplace_kernel, 
+        gamma0=gamma0, n_iter=n_iter, search_n=search_n, search_loexp=search_loexp, train_split=split)
+    model.fit(Xtr, Ytr)
+
+    iters = list(range(n_iter))
+    refit_mses = []
+    wiggle_mses = []
+    gamma_hist = []
+    for it in iters:
+        Ypr = model.model_hist[it].predict(Xte)
+        refit_mses.append(mean_squared_error(Yte, Ypr))
+        wiggle_mses.append(model.model_hist[it].best_mse)
+    
+    np.savez(RES_FILE, refit=refit_mses, wiggle=wiggle_mses, gamma_hist=model.gamma_hist)
+
+def figure(show=False):
+    iters = list(range(n_iter))
+    results = np.load(RES_FILE)
+
+    fig,axs = plt.subplots(1, 2)
+    fig.set_size_inches(9,3)
+    axs[0].set_title('MSE history')
+    axs[0].plot(iters, results['refit'], label='refit with best bandwidth')
+    axs[0].plot(iters, results['wiggle'], label='with best wiggled bandwidth')
+    axs[0].legend()
+    axs[1].set_title('$\gamma$ history')
+    axs[1].plot(iters, results['gamma_hist'])
+
+    if show:
+        plt.show()
+    else:
+        fig.savefig(FIG_FILE)

### 2023-01-17 code/mnist.py

@@ -1,43 +1,0 @@
-import os
-import numpy as np
-from sklearn.datasets import fetch_openml
-
-MNIST_PATH = 'build/mnist_784.npz'
-
-class MnistBinaryDataset:
-
-    def __init__(self, seed):
-        if not os.path.exists(MNIST_PATH):
-            mnist = fetch_openml('mnist_784', parser='auto')
-            self.allX = mnist.data.to_numpy()
-            self.allY = mnist.target.astype(int).to_numpy()
-            np.savez(MNIST_PATH, X=self.allX, Y=self.allY)
-        else:
-            archive = np.load(MNIST_PATH)
-            self.allX = archive['X']
-            self.allY = archive['Y']
-        self.dim = self.allX.shape[1]
-        self.rng = np.random.default_rng(seed)
-
-    def sample(self, digitA, digitB, split):
-        assert(np.all(np.array(split) > 0))
-        relevant = np.logical_or(self.allY == digitA, self.allY == digitB)
-        N = np.sum(split)
-        assert(np.sum(relevant) >= N)
-
-        perm = self.rng.permutation(N)
-        X = self.allX[relevant][:N][perm]
-        Y = self.allY[relevant][:N][perm].reshape(-1,1)
-        Y[Y == digitA] = -1
-        Y[Y == digitB] = 1
-
-        curPos = 0
-        r = []
-        for p in split:
-            r.append(X[curPos:curPos+p])
-            r.append(Y[curPos:curPos+p])
-            curPos += p
-        return r
-
-if __name__ == '__main__':
-    MnistBinaryDataset(0)

### 2023-01-17 code/shared.py

@@ -1,0 +1,149 @@
+import os
+import time
+import numpy as np
+from scipy.spatial import distance_matrix
+from sklearn.datasets import fetch_openml
+from sklearn.metrics import mean_squared_error
+from sklearn.kernel_ridge import KernelRidge
+import numpy as np
+
+
+
+MNIST_PATH = 'build/mnist_784.npz'
+
+class MnistBinaryDataset:
+
+    def __init__(self, seed):
+        if not os.path.exists(MNIST_PATH):
+            mnist = fetch_openml('mnist_784', parser='auto')
+            self.allX = mnist.data.to_numpy()
+            self.allY = mnist.target.astype(int).to_numpy()
+            np.savez(MNIST_PATH, X=self.allX, Y=self.allY)
+        else:
+            archive = np.load(MNIST_PATH)
+            self.allX = archive['X']
+            self.allY = archive['Y']
+        self.dim = self.allX.shape[1]
+        self.rng = np.random.default_rng(seed)
+
+    def sample(self, digitA, digitB, split):
+        assert(np.all(np.array(split) > 0))
+        relevant = np.logical_or(self.allY == digitA, self.allY == digitB)
+        N = np.sum(split)
+        assert(np.sum(relevant) >= N)
+
+        perm = self.rng.permutation(N)
+        X = self.allX[relevant][:N][perm]
+        Y = self.allY[relevant][:N][perm].reshape(-1,1)
+        Y[Y == digitA] = -1
+        Y[Y == digitB] = 1
+
+        curPos = 0
+        r = []
+        for p in split:
+            r.append(X[curPos:curPos+p])
+            r.append(Y[curPos:curPos+p])
+            curPos += p
+        return r
+
+
+
+def laplace_kernel(gamma,X0,X1=None):
+    if X1 is None: X1 = X0
+    D = distance_matrix(X0, X1, p=2)
+    return np.exp(-gamma * D)
+
+def gaussian_kernel(gamma,X0,X1=None):
+    if X1 is None: X1 = X0
+    D = distance_matrix(X0, X1, p=2)
+    return np.exp(-gamma * D**2)
+
+
+
+class WiggledScaleKernelLS:
+
+    def __init__(self, kernel_func, base_gamma, wiggled_gamma=None):
+        self.kernel_func = kernel_func
+        self.base_gamma = base_gamma
[…]

### 2023-01-17 experiments/__init__.py
fix python packaging for this repo

@@ -1,0 +1,1 @@
+# nothing here... makes this a package

### 2023-01-17 experiments/exp_iterative_algo.py

@@ -1,7 +1,7 @@
 import numpy as np
 from sklearn.metrics import mean_squared_error
 from matplotlib import pyplot as plt
-from code.shared import MnistBinaryDataset, laplace_kernel, IteratedWiggleSearchScaleKernelLS
+from experiments.shared import MnistBinaryDataset, laplace_kernel, IteratedWiggleSearchScaleKernelLS
 
 RES_FILE = 'build/results/iterative_algo.npz'
 FIG_FILE = 'build/figures/iterative_algo.png'

### 2023-01-17 experiments/exp_iterative_algo.py
improve makefile by decoupling computation and presentation

@@ -1,51 +1,0 @@
-import numpy as np
-from sklearn.metrics import mean_squared_error
-from matplotlib import pyplot as plt
-from experiments.shared import MnistBinaryDataset, laplace_kernel, IteratedWiggleSearchScaleKernelLS
-
-RES_FILE = 'build/results/iterative_algo.npz'
-FIG_FILE = 'build/figures/iterative_algo.png'
-
-Ntr = 200
-Nte = 500
-Xtr,Ytr, Xte,Yte = MnistBinaryDataset(3281).sample(4, 9, [Ntr, Ntr])
-
-gamma0 = 1/784
-n_iter = 8
-search_n = 5
-search_loexp = 2
-split = 0.85
-
-def compute():
-    model = IteratedWiggleSearchScaleKernelLS(kernel_func=laplace_kernel, 
-        gamma0=gamma0, n_iter=n_iter, search_n=search_n, search_loexp=search_loexp, train_split=split)
-    model.fit(Xtr, Ytr)
-
-    iters = list(range(n_iter))
-    refit_mses = []
-    wiggle_mses = []
-    gamma_hist = []
-    for it in iters:
-        Ypr = model.model_hist[it].predict(Xte)
-        refit_mses.append(mean_squared_error(Yte, Ypr))
-        wiggle_mses.append(model.model_hist[it].best_mse)
-    
-    np.savez(RES_FILE, refit=refit_mses, wiggle=wiggle_mses, gamma_hist=model.gamma_hist)
-
-def figure(show=False):
-    iters = list(range(n_iter))
-    results = np.load(RES_FILE)
-
-    fig,axs = plt.subplots(1, 2)
-    fig.set_size_inches(9,3)
-    axs[0].set_title('MSE history')
-    axs[0].plot(iters, results['refit'], label='refit with best bandwidth')
-    axs[0].plot(iters, results['wiggle'], label='with best wiggled bandwidth')
-    axs[0].legend()
-    axs[1].set_title('$\gamma$ history')
-    axs[1].plot(iters, results['gamma_hist'])
-
-    if show:
-        plt.show()
-    else:
-        fig.savefig(FIG_FILE)

### 2023-01-17 experiments/iterative_algo_compute.py

@@ -1,0 +1,29 @@
+import numpy as np
+from sklearn.metrics import mean_squared_error
+from shared import MnistBinaryDataset, laplace_kernel, IteratedWiggleSearchScaleKernelLS, get_result_and_figure_path
+
+if __name__ == '__main__':
+    Ntr = 2000
+    Nte = 2000
+    Xtr,Ytr, Xte,Yte = MnistBinaryDataset(3281).sample(4, 9, [Ntr, Ntr])
+    gamma0 = 1/784
+    n_iter = 10
+    search_n = 20
+    search_loexp = 2
+    split = 0.85
+
+    model = IteratedWiggleSearchScaleKernelLS(kernel_func=laplace_kernel, 
+        gamma0=gamma0, n_iter=n_iter, search_n=search_n, search_loexp=search_loexp, train_split=split)
+    model.fit(Xtr, Ytr)
+
+    iters = list(range(n_iter))
+    refit_mses = []
+    wiggle_mses = []
+    gamma_hist = []
+    for it in iters:
+        Ypr = model.model_hist[it].predict(Xte)
+        refit_mses.append(mean_squared_error(Yte, Ypr))
+        wiggle_mses.append(model.model_hist[it].best_mse)
+    
+    resFile,_ = get_result_and_figure_path()
+    np.savez(resFile, refit=refit_mses, wiggle=wiggle_mses, gamma_hist=model.gamma_hist)

### 2023-01-17 experiments/iterative_algo_present.py

@@ -1,0 +1,20 @@
+from shared import get_result_and_figure_path
+import numpy as np
+from matplotlib import pyplot as plt
+
+if __name__ == '__main__':
+    resFile, figFile = get_result_and_figure_path()
+    results = np.load(resFile)
+    n_iter = len(results['refit'])
+    iters = list(range(n_iter))
+
+    fig,axs = plt.subplots(1, 2)
+    fig.set_size_inches(9,3)
+    axs[0].set_title('MSE history')
+    axs[0].plot(iters, results['refit'], label='refit with best bandwidth')
+    axs[0].plot(iters, results['wiggle'], label='with best wiggled bandwidth')
+    axs[0].legend()
+    axs[1].set_title('$\gamma$ history')
+    axs[1].plot(iters, results['gamma_hist'])
+
+    fig.savefig(figFile)

### 2023-01-17 experiments/shared.py

@@ -1,3 +1,4 @@
+import sys
 import os
 import time
 import numpy as np
@@ -6,6 +7,15 @@ from sklearn.datasets import fetch_openml
 from sklearn.metrics import mean_squared_error
 from sklearn.kernel_ridge import KernelRidge
 import numpy as np
+
+
+
+def get_result_and_figure_path():
+    script = sys.argv[0]
+    basename = os.path.splitext(os.path.basename(script))[0]\
+        .replace('_compute', '').replace('_present', '')
+    return ( 'build/results/' + basename + '.npz', 
+        'build/figures/' + basename + '.png' )
 
 
 

### 2023-01-17 experiments/iterative_algo_compute.py
improve iterative and add split experiment

@@ -26,4 +26,6 @@ if __name__ == '__main__':
         wiggle_mses.append(model.model_hist[it].best_mse)
     
     resFile,_ = get_result_and_figure_path()
-    np.savez(resFile, refit=refit_mses, wiggle=wiggle_mses, gamma_hist=model.gamma_hist)
+    np.savez(resFile, 
+        refit=refit_mses, wiggle=wiggle_mses, 
+        gamma_hist=model.gamma_hist, fit_times=model.fit_times)

### 2023-01-17 experiments/iterative_algo_present.py

@@ -9,12 +9,17 @@ if __name__ == '__main__':
     iters = list(range(n_iter))
 
     fig,axs = plt.subplots(1, 2)
-    fig.set_size_inches(9,3)
+    fig.set_size_inches(9,4.5)
+    fig.suptitle('Average fit time: {:.2f} seconds'.format(np.mean(np.array(results['fit_times']))))
     axs[0].set_title('MSE history')
     axs[0].plot(iters, results['refit'], label='refit with best bandwidth')
     axs[0].plot(iters, results['wiggle'], label='with best wiggled bandwidth')
+    axs[0].set_xlabel('iteration')
+    axs[0].set_ylabel('MSE')
     axs[0].legend()
-    axs[1].set_title('$\gamma$ history')
+    axs[1].set_title('bandwidth history')
     axs[1].plot(iters, results['gamma_hist'])
+    axs[0].set_xlabel('iteration')
+    axs[0].set_ylabel('$\gamma$')
 
     fig.savefig(figFile)

### 2023-01-17 experiments/shared.py

@@ -79,8 +79,10 @@ class WiggledScaleKernelLS:
         self.model = KernelRidge(alpha=0, gamma=base_gamma, kernel='precomputed')
 
     def fit(self, Xtr, Ytr):
+        startTime = time.time()
         Ktr = self.kernel_func(self.base_gamma, Xtr)
         self.model.fit(Ktr, Ytr)
+        self.fit_time = time.time() - startTime
         self.Xtr = Xtr
         return self
 
@@ -107,14 +109,18 @@ class WiggleSearchScaleKernelLS:
         Ytr0 = Ytr[:Ntr0]
         Xtr1 = Xtr[Ntr0:]
         Ytr1 = Ytr[Ntr0:]
+
         self.model.fit(Xtr0, Ytr0)
+        self.fit_time = self.model.fit_time
 
         mses = []
+        startTime = time.time()
         for gamma in self.search_gammas:
             self.model.wiggled_gamma = gamma
             Ypr1 = self.model.predict(Xtr1)
             mses.append(mean_squared_error(Ytr1, Ypr1))
         i = np.argmin(np.array(mses))
+        self.wiggle_time = time.time() - startTime
         self.best_gamma = self.search_gammas[i]
         self.model.wiggled_gamma = self.best_gamma
         self.best_mse = mses[i]
@@ -153,7 +159,7 @@ class IteratedWiggleSearchScaleKernelLS:
             cur_gamma = self.model_hist[-1].best_gamma
             self.gamma_hist.append(cur_gamma)
             self.fit_times.append(time.time() - startTime)
-            print('iteration {}/{} done in {} seconds'.format(it + 1, self.n_iter, self.fit_times[-1]))
+            print('iteration {0}/{1} done in {2:.2f} seconds'.format(it + 1, self.n_iter, self.fit_times[-1]))
 
     def predict(self, Xte):
         self.model_hist[-1].predict(Xte)

### 2023-01-17 experiments/split_compute.py

@@ -1,0 +1,41 @@
+import time
+import numpy as np
+from sklearn.metrics import mean_squared_error
+from shared import MnistBinaryDataset, laplace_kernel, WiggledScaleKernelLS, WiggleSearchScaleKernelLS, get_search_gammas, get_result_and_figure_path
+
+if __name__ == '__main__':
+    Ntr = 1000
+    Nte = 1000
+    Xtr,Ytr, Xte,Yte = MnistBinaryDataset(48888).sample(4, 9, [Ntr, Ntr])
+    splits = np.linspace(0.3, 0.95, num=20)
+    gamma0 = 1/784
+    search_gammas = get_search_gammas(gamma0, 2, 15)
+    best_gammas = []
+    test_mses = []
+    sample_mses = []
+    wiggle_mses = []
+    fit_times = []
+    wiggle_times = []
+
+    baseline = WiggledScaleKernelLS(laplace_kernel, gamma0).fit(Xtr,Ytr)
+    baseline_mse = mean_squared_error(baseline.predict(Xte), Yte)
+    baseline_sample_mse = mean_squared_error(baseline.predict(Xtr), Ytr)
+    print('fit baseline model in {:.2f} seconds'.format(baseline.fit_time))
+
+    for split in splits:
+        model = WiggleSearchScaleKernelLS(laplace_kernel, gamma0, search_gammas, split)
+        model.fit(Xtr, Ytr)
+        fit_times.append(model.fit_time)
+        wiggle_times.append(model.wiggle_time)
+        best_gammas.append(model.best_gamma)
+        wiggle_mses.append(model.best_mse)
+        test_mses.append(mean_squared_error(Yte, model.predict(Xte)))
+        sample_mses.append(mean_squared_error(Ytr, model.predict(Xtr)))
+        print('finished train split at {:.1f}% with total training time of {:.2f} seconds'.format(split*100.0, fit_times[-1] + wiggle_times[-1]))
+
+    resFile,_ = get_result_and_figure_path()
+    np.savez(resFile, 
+        baseline=baseline_mse, baseline_sample=baseline_sample_mse,
+        splits=splits, best_gammas=best_gammas,
+        test_mses=test_mses, sample_mses=sample_mses, wiggle_mses=wiggle_mses,
+        fit_times=fit_times, wiggle_times=wiggle_times, baseline_fit_time=baseline.fit_time)

### 2023-01-17 experiments/split_present.py

@@ -1,0 +1,45 @@
+from shared import get_result_and_figure_path
+import numpy as np
+from matplotlib import pyplot as plt
+
+if __name__ == '__main__':
+    resFile, figFile = get_result_and_figure_path()
+    results = np.load(resFile)
+    splits = results['splits']
+
+    fig,axs = plt.subplots(2, 2)
+    fig.set_size_inches(10,10)
+
+    axs[0,0].set_title('sample MSE')
+    axs[0,0].plot([splits[0],splits[-1]], 2*[results['baseline_sample']], label='baseline', c='black')
+    axs[0,0].plot(splits, results['sample_mses'], label='wiggle search')
+    axs[0,0].set_xlabel('train split')
+    axs[0,0].set_ylabel('MSE')
+    axs[0,0].legend()
+
+    axs[0,1].set_title('test MSE')
+    axs[0,1].plot([splits[0],splits[-1]], 2*[results['baseline']], label='baseline', c='black')
+    axs[0,1].plot(splits, results['test_mses'], label='wiggle search on test')
+    axs[0,1].plot(splits, results['wiggle_mses'], label='wiggle search on internal test')
+    axs[0,1].set_xlabel('train split')
+    axs[0,1].set_ylabel('MSE')
+    axs[0,1].legend()
+
+    axs[1,0].set_title('chosen bandwidths')
+    axs[1,0].plot(splits, results['best_gammas'])
+    axs[1,0].set_yscale('log')
+    axs[1,0].set_xlabel('train split')
+    axs[1,0].set_ylabel('$\gamma$')
+
+    axs[1,1].set_title('fit times')
+    axs[1,1].plot([splits[0],splits[-1]], 2*[results['baseline_fit_time']], label='baseline', c='black')
+    total = np.array(results['wiggle_times']) + np.array(results['fit_times'])
+    wiggle = np.array(results['wiggle_times'])
+    axs[1,1].fill_between(splits, total, label='least squares')
+    axs[1,1].fill_between(splits, wiggle, label='wiggle')
+    axs[1,1].plot(splits, total)
+    axs[1,1].set_xlabel('train split')
+    axs[1,1].set_ylabel('seconds')
+    axs[1,1].legend()
+
+    fig.savefig(figFile)

### 2023-01-17 paper.tex
finalize table of contents and include ready figures

@@
 \usepackage[nopatch=eqnum]{microtype}
+\usepackage{graphicx}
 \usepackage{cleveref}
+
+\graphicspath{{build/figures/}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% kernel ridgeless regression
 Let $\mathbf X \in \R^{n \times d}$ be a matrix of $n$ training examples
@@
 
+% new regimes of statistical learning
 Recently, there was a new wave of interest into kernels sparked by the 
@@
 
+% focus on kernel methods
 In particular, there are many indications that kernel learning methods 
@@
 
-These reasons have motivated our choice to consider here the extrema of 
-the spectrum of \emph{Matérn kernels}: a family of kernels whose associated
-RKHSs are exactly Sobolev spaces with real smoothness index.
-The least smooth member of the family, the \emph{Laplacian kernel}, is 
-defined by
-\[ k^L_\gamma(x,x') = \exp(-\gamma \| x - x' \|) \]
-and the smoothest member of it is the \emph{Gaussian kernel}
-\[ k^G_\gamma(x,x') = \exp(-\gamma \| x - x' \|^2). \]
+% choice of kernel and what we will do
+These considerations have also sparked new interest in the other direction:
+due to all the analogies between kernels and neural nets researchers have
+started to scale up kernel methods to deal with big data.
+Interesting targets have been kernels coming from the Sobolev-Hilbert spaces
+$H^s(\R^d)$ with real smoothness index: the \emph{Matérn kernels}.
+This family's smoothest member is the Gaussian kernel, but we will (due to 
+the motivations above) consider only its least smooth member: the Laplacian
+kernel
+\[ k_\gamma(x,x') = \exp(-\gamma \| x - x' \|). \]
+The only hyper-parameter to adjust is the bandwidth $\gamma$ and in this 
+paper we investigate a new principle on how to both optimize it in 
+large-scale contexts and use it for regularization purposes.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Wiggling as a regularization technique}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Wiggling}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% what i mean by wiggling
+
+% not interpolating anymore
+
+% computational complexity and algorithmics
+
+% questions
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Iterative wiggling to find the best bandwidth}
+\subsection{Experiment: Does it compare to Tikhonov?}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions and directions}
+\subsection{Experiment: Where to split the training set?}
+
+\begin{figure}[htp]
+    \centering
+    \includegraphics[width=\textwidth]{split}
+    \caption{split}
+    \label{fig:split}
+\end{figure}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experiment: Iterative wiggling as cheap bandwidth optimization}
+
+\begin{figure}[htp]
+    \centering
+    \includegraphics[width=\textwidth]{iterative_algo}
+    \caption{split}
+    \label{fig:iterative_algo}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions and future directions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{thebibliography}{20}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2023-01-17 experiments/shared.py
drastically improve efficiency by caching the distance matrix in wiggle searches

@@ -58,15 +58,10 @@ class MnistBinaryDataset:
 
 
 
-def laplace_kernel(gamma,X0,X1=None):
-    if X1 is None: X1 = X0
-    D = distance_matrix(X0, X1, p=2)
+def laplace_kernel(gamma,X0=None,X1=None,D=None):
+    if D is None and X1 is None: X1 = X0
+    if D is None: D = distance_matrix(X0, X1, p=2)
     return np.exp(-gamma * D)
-
-def gaussian_kernel(gamma,X0,X1=None):
-    if X1 is None: X1 = X0
-    D = distance_matrix(X0, X1, p=2)
-    return np.exp(-gamma * D**2)
 
 
 
@@ -77,6 +72,8 @@ class WiggledScaleKernelLS:
         self.base_gamma = base_gamma
         self.wiggled_gamma = wiggled_gamma if wiggled_gamma is not None else base_gamma
         self.model = KernelRidge(alpha=0, gamma=base_gamma, kernel='precomputed')
+        self.Xtr = None
+        self.distmatrix = None
 
     def fit(self, Xtr, Ytr):
         startTime = time.time()
@@ -85,9 +82,16 @@ class WiggledScaleKernelLS:
         self.fit_time = time.time() - startTime
         self.Xtr = Xtr
         return self
+    
+    def precompute_eval_distmatrix(self, Xte):
+        assert(self.Xtr is not None)
+        self.distmatrix = distance_matrix(Xte, self.Xtr, p=2)
 
-    def predict(self, Xte):
-        Kte = self.kernel_func(self.wiggled_gamma, Xte, self.Xtr)
+    def predict(self, Xte=None):
+        if Xte is None:
+            Kte = self.kernel_func(self.wiggled_gamma, D=self.distmatrix)
+        else:
+            Kte = self.kernel_func(self.wiggled_gamma, X0=Xte, X1=self.Xtr)
         return self.model.predict(Kte)
 
 
@@ -112,12 +116,13 @@ class WiggleSearchScaleKernelLS:
 
         self.model.fit(Xtr0, Ytr0)
         self.fit_time = self.model.fit_time
+        self.model.precompute_eval_distmatrix(Xtr1)
 
         mses = []
         startTime = time.time()
         for gamma in self.search_gammas:
             self.model.wiggled_gamma = gamma
-            Ypr1 = self.model.predict(Xtr1)
+            Ypr1 = self.model.predict()
[…]

### 2023-01-18 experiments/iterative_algo_compute.py
fix copy-paste error in train test sizes

@@ -5,7 +5,7 @@ from shared import MnistBinaryDataset, laplace_kernel, IteratedWiggleSearchScale
 if __name__ == '__main__':
     Ntr = 2000
     Nte = 2000
-    Xtr,Ytr, Xte,Yte = MnistBinaryDataset(3281).sample(4, 9, [Ntr, Ntr])
+    Xtr,Ytr, Xte,Yte = MnistBinaryDataset(3281).sample(4, 9, [Ntr, Nte])
     gamma0 = 1/784
     n_iter = 10
     search_n = 20

### 2023-01-18 experiments/split_compute.py

@@ -6,7 +6,7 @@ from shared import MnistBinaryDataset, laplace_kernel, WiggledScaleKernelLS, Wig
 if __name__ == '__main__':
     Ntr = 1000
     Nte = 1000
-    Xtr,Ytr, Xte,Yte = MnistBinaryDataset(48888).sample(4, 9, [Ntr, Ntr])
+    Xtr,Ytr, Xte,Yte = MnistBinaryDataset(48888).sample(4, 9, [Ntr, Nte])
     splits = np.linspace(0.3, 0.95, num=20)
     gamma0 = 1/784
     search_gammas = get_search_gammas(gamma0, 2, 15)

### 2023-01-18 experiments/shared.py
add tikhonov parameter

@@ -67,11 +67,11 @@ def laplace_kernel(gamma,X0=None,X1=None,D=None):
 
 class WiggledScaleKernelLS:
 
-    def __init__(self, kernel_func, base_gamma, wiggled_gamma=None):
+    def __init__(self, kernel_func, base_gamma, wiggled_gamma=None, alpha=0):
         self.kernel_func = kernel_func
         self.base_gamma = base_gamma
         self.wiggled_gamma = wiggled_gamma if wiggled_gamma is not None else base_gamma
-        self.model = KernelRidge(alpha=0, gamma=base_gamma, kernel='precomputed')
+        self.model = KernelRidge(alpha=alpha, gamma=base_gamma, kernel='precomputed')
         self.Xtr = None
         self.distmatrix = None
 
@@ -98,12 +98,12 @@ class WiggledScaleKernelLS:
 
 class WiggleSearchScaleKernelLS:
 
-    def __init__(self, kernel_func, gamma0, search_gammas, split):
+    def __init__(self, kernel_func, gamma0, search_gammas, split, alpha=0):
         self.kernel_func = kernel_func
         self.gamma0 = gamma0
         self.search_gammas = search_gammas
         self.split = split
-        self.model = WiggledScaleKernelLS(kernel_func=kernel_func, base_gamma=gamma0)
+        self.model = WiggledScaleKernelLS(kernel_func=kernel_func, base_gamma=gamma0, alpha=alpha)
         self.best_gamma = None
 
     def fit(self, Xtr, Ytr):

### 2023-01-18 experiments/tikhonov_compute.py
design comparison with tikhonov

@@ -1,0 +1,39 @@
+import time
+import numpy as np
+from sklearn.metrics import mean_squared_error
+from shared import MnistBinaryDataset, laplace_kernel, WiggledScaleKernelLS, WiggleSearchScaleKernelLS, get_search_gammas, get_result_and_figure_path
+
+if __name__ == '__main__':
+    Ntr = 1000
+    Nte = 1000
+    Xtr,Ytr, Xte,Yte = MnistBinaryDataset(98).sample(4, 9, [Ntr, Nte])
+
+    gamma0 = 1/784
+    alphas = np.logspace(-5, 1, num=15)
+
+    baseline = WiggledScaleKernelLS(laplace_kernel, base_gamma=gamma0)
+    baseline.fit(Xtr, Ytr)
+    baseline_mse = mean_squared_error(Yte, baseline.predict(Xte))
+    wiggled = WiggleSearchScaleKernelLS(laplace_kernel, gamma0, get_search_gammas(gamma0, 2, 20), 0.85)
+    wiggled.fit(Xtr, Ytr)
+    wiggled_mse = mean_squared_error(Yte, wiggled.predict(Xte))
+
+    tikhonov_mses = []
+    wiggled_tikhonov_mses = []
+    for i in range(alphas.size):
+        alpha = alphas[i]
+
+        startTime = time.time()
+        tikh = WiggledScaleKernelLS(laplace_kernel, base_gamma=gamma0, alpha=alpha)
+        tikh.fit(Xtr, Ytr)
+        tikhonov_mses.append(mean_squared_error(Yte, tikh.predict(Xte)))
+
+        wtikh = WiggleSearchScaleKernelLS(laplace_kernel, gamma0, get_search_gammas(gamma0, 2, 20), 0.85, alpha=alpha)
+        wtikh.fit(Xtr, Ytr)
+        wiggled_tikhonov_mses.append(mean_squared_error(Yte, wtikh.predict(Xte)))
+
+        print('finished alpha {}/{} in {:.2f} seconds'.format(i+1, alphas.size, time.time() - startTime))
+
+    resFile,_ = get_result_and_figure_path()
+    np.savez(resFile, alphas=alphas, baseline_mse=baseline_mse, wiggled_mse=wiggled_mse, 
+        tikhonov_mses=tikhonov_mses, wiggled_tikhonov_mses=wiggled_tikhonov_mses)

### 2023-01-18 experiments/tikhonov_present.py

@@ -1,0 +1,18 @@
+from shared import get_result_and_figure_path
+import numpy as np
+from matplotlib import pyplot as plt
+
+if __name__ == '__main__':
+    resFile, figFile = get_result_and_figure_path()
+    results = np.load(resFile)
+    alphas = results['alphas']
+
+    plt.plot([alphas[0], alphas[-1]], 2*[results['baseline_mse']], label='baseline')
+    plt.plot(alphas, results['tikhonov_mses'], label='tikhonov')
+    plt.plot([alphas[0], alphas[-1]], 2*[results['wiggled_mse']], label='wiggled')
+    plt.plot(alphas, results['wiggled_tikhonov_mses'], label='wiggled tikhonov')
+    plt.xscale('log')
+    plt.xlabel('$\\alpha$')
+    plt.ylabel('MSE')
+    plt.legend()
+    plt.savefig(figFile)

### 2023-01-28 paper.tex
finish intro and questions

@@
 \usepackage{cleveref}
+\usepackage{mathtools}
+
+\setlength{\textwidth}{\paperwidth}
+\addtolength{\textwidth}{-2in}
+\calclayout
 
@@
 \newcommand{\R}{\mathbb{R}}
+\DeclareMathOperator{\Tr}{Tr}
+\DeclareMathOperator{\argmin}{argmin}
 
-\title{Wiggling the bandwidth in ridgeless regression 
-        with scale-dependent kernels}
+\title{Wiggling the bandwidth}
 \author{Luca Wellmeier}
@@
     The provided experimental results indicate that \textit{wiggling}
-    the bandwidth in this way can generalize and be a cheap alternative to
-    the classical Tikhonov square-norm penality.
+    the bandwidth in this way can provide generalization and potentially 
+    be a cheap alternative to the classical Tikhonov square-norm penality.
     Finally, we propose and test a new iterative method for optimizing the 
-    bandwidth parameter by employing wiggling in a local search.
+    bandwidth parameter by employing wiggling in a cheap local parameter 
+    search.
 \end{abstract}
@@
 with low effective dimension (\cite{benignoverfitting})
-and properly chosen smoothness of the kernel w.r.t. the nature of the data 
+and properly chosen curvature of the kernel w.r.t. the nature of the data 
 (\cite{justinterpolate}).
@@
 due to all the analogies between kernels and neural nets researchers have
-started to scale up kernel methods to deal with big data.
+started to scale up kernel methods to deal with big data 
+(see for instance \cite{falkon}).
 Interesting targets have been kernels coming from the Sobolev-Hilbert spaces
 $H^s(\R^d)$ with real smoothness index: the \emph{Matérn kernels}.
-This family's smoothest member is the Gaussian kernel, but we will (due to 
-the motivations above) consider only its least smooth member: the Laplacian
-kernel
-\[ k_\gamma(x,x') = \exp(-\gamma \| x - x' \|). \]
-The only hyper-parameter to adjust is the bandwidth $\gamma$ and in this 
-paper we investigate a new principle on how to both optimize it in 
-large-scale contexts and use it for regularization purposes.
+This family's smoothness limit member is the Gaussian kernel, but we will 
+(due to the motivations above) 
+consider only the well known the Laplacian kernel
+\[ k_\gamma(x,x') = \exp(-\gamma \| x - x' \|) \]
+which belongs to the less smooth part of the spectrum.
+The only hyperparameter to adjust is the bandwidth $\gamma$ and in this 
+paper we investigate experimentally a new principle on how to both optimize 
+it in large-scale contexts and use it for regularization purposes.
+
+The full code of the experiments can be found 
+on GitHub\footnote{\url{https://github.com/lucawellmeier/wiggle-bandwidth}}.
 
@@
 
+% notation and basics
+As announced, let us fix the Laplacian kernel function
+\[ k_\gamma(x,x') = \exp\left(-\frac{\| x - x' \|}{\gamma} \right) \]
+with bandwidth $\gamma > 0$.
+The convention here is that smaller bandwidth actually decreases the 
+area under the curve $x \mapsto k_0(x) \coloneqq k(0,x)$ and vice-versa.
+This whole note is based on the observation that the kernel ridgeless 
+estimators use the bandwidth twice in distinct places:
+once in the computation of the \emph{dual coefficients}
+\[ c_i = (\mathbf K_\gamma^\dagger Y)_i \]
+where the kernel matrix depends on it and then
+in the evaluation part
+\[ \sum_i c_i k_\gamma(x,x_i) \]
+where the kernel is computed to check similarities between sample 
+points $x_i$ and unseen points $x$.
+Note that the dependence of the dual coefficients on the bandwidth is 
+not even guaranteed to be continuous and generally hard to understand
+due to the appearing pseudo-inverse.
+However, the evaluation part offers a different perspective, if we 
+fix the dual coefficients $c_i$ and an unseen point $x$, then
+the relationship
+\[ \gamma \mapsto \sum_{i=1}^n c_i k_\gamma(x,x_i) \]
+is smooth: indeed, $\gamma$ appears only as a factor to the non-smooth
+norm in the exponential.
+
 % what i mean by wiggling
+Thus, we propose the new \emph{bandwidth-wiggled} estimators
+\[ \hat f_{\gamma_0, \gamma}(x) \coloneqq 
+    \sum_{i=1}^n (K_{\gamma_0} Y)^\dagger k_\gamma(x, x_i). \]
+The pseudo-inverse is computed for a given base bandwidth $\gamma_0$
+but the evaluation part is performed with a different bandwidth $\gamma$.
+Due to the regularity argument above, we hypothesize that 
+changing $\gamma_0 \to \gamma$ in the evaluation part is well-behaved.
+The next chapters are devoted to experimentally verifying this, but 
+also to explore possible new methods based on this estimator.
 
-% not interpolating anymore
+% wiggle search
+The first method that we introduce here is referred to as 
+\emph{wiggle search}.
+Let $\Tr$ denote the training set with $n$ samples.
+Split it into two non-empty parts: $\Tr = \Tr_1 \sqcup \Tr_2$
+and train the a standard Laplacian ridgeless estimator only on
+$\Tr_1$ with an initial guessed bandwidth $\gamma_0$.
+Then extract the dual coefficients $c_i^{\gamma_0}$ from there 
+and use them to define the family predictors 
+\[ \hat f_{\gamma_0,\gamma} 
+    \coloneqq \sum_{i \in \Tr_1} c_i^{\gamma_0} k_\gamma(\cdot, x_i) \]
+for each $\gamma$ in a finite set $\Gamma \subset \R_{> 0}$.
+Then select $\gamma^*$ such that 
+\[ \gamma^* = \argmin_{\gamma \in \Gamma} 
+    \hat R_{\Tr_2} \left(\hat f_{\gamma_0,\gamma}\right), \]
+i.e. the associated wiggled estimator minimizes the empricial mean
+squared error on the held-out part of the sample.
+This algorithm is more efficient in comparison to computing the 
+standard estimators:
+The least squares computation to determine the dual coefficients is
+cubic in the sample size, so as a first improvement, we reduce this 
+time effectively by splitting the sample.
+Moreover, existing techniques for optimizing the bandwidth
+(i.e. grid search, cross-validation, ...) test one parameter value 
+per least squares computation.
+Here, instead, potentially thousands of parameters are tested with 
+on single least squares assuming that $\hat f_{\gamma_0,\gamma}$ 
+is representative of the actual standard estimator 
+$\hat f_{\gamma,\gamma}$.
 
-% computational complexity and algorithmics
+% computational trick
+Note that the comparison of the different bandwidths can be computed 
+in a highly vectorized way:
+Let $\mathbf D$ denote the distance matrix of 2-norms between $\Tr_1$ and 
+$\Tr_2$.
+Once it is computed, we can test a new wiggled estimator 
+for \emph{any} other bandwidth $\gamma$ by running
+the vectorized computation
+\[ \big( \mathbf c^{\gamma_0} \exp(-(1/\gamma) \mathbf D) \big)^T \]
+yielding the new predictions of the inputs in $\Tr_2$.
+Thus, the big hard part of the computation, i.e. finding $\mathbf D$,
+needs to be done only once for size of $\Gamma$.
+
+% iterative search
+If wiggle search performs well and compares to the standard estimator 
+with the optimal wiggled bandwidth (as we will see in the experiments), 
+we can start considering an \emph{itertated wiggle search}:
+run a wiggle search, find the optimal bandwidth, recompute the 
+least squares with that bandwidth on $\Tr_1$ and repeat.
+That will make up our final experiment in the paper.
 
 % questions
+Summing up, here is the questions we aim to provide experimental 
+evidence for:
+\begin{enumerate}
+    \item How close are wiggled estimators and standard estimators with 
+        the wiggle bandwidth? 
+        What is the radius of $\gamma$'s around $\gamma_0$ such that one 
+        can move knowledge from one to the other? 
+        How to choose the wiggle range $\Gamma$?
+    \item What would be a good point to split the training set in the 
+        context of wiggle search?
+        When does wiggle search provide an advantage over standard 
+        estimators (or if at all)?
+    \item Given that we move away from the interpolating state 
+        (in a smooth and controlled manner) does wiggling compare to 
+        Tikhonov?
+        What happens with the norm/complexitie of a wiggled estimator
+        compared to the standard one?
+    \item Does an iterated wiggle search find the best bandwidth?
+        Is it less expensive than classical model selection techniques?
+\end{enumerate}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experiment: How close are wiggled and standard estimators?}
 
@@
 \subsection{Experiment: Does it compare to Tikhonov?}
+
+\begin{figure}[htp]
+    \centering
+    \includegraphics[width=\textwidth]{tikhonov}
+    \caption{tikhonov}
+    \label{fig:tikhonov}
+\end{figure}
 
@@
     \includegraphics[width=\textwidth]{iterative_algo}
-    \caption{split}
+    \caption{iterative algo}
     \label{fig:iterative_algo}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions and future directions}
+\section{Conclusions and next steps}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Finally, I propose a list of todos and questions for further research on 
+the method.
+\begin{itemize}
+    \item How far is the reach of wiggling being indicative about the refit
+        performance? That is, if I fix an initial bandwidth $\gamma_0$, 
+        what is the range of $\gamma$ such that the performance of the 
+        estimator $f_{\gamma_0, \gamma}$ can predict that of the estimator
+        $f_\gamma$? In small logarithmic local ranges our experiments have
+        shown strong similaritity of the risk curves but how far can this be 
+        streched? Here, theoretical guarantees are needed to make the method
+        pratically usable.
+    \item The MNIST dataset have been shown in \cite{justinterpolate} to 
+        be predicted best by ridgeless kernel estimators.
+        How do the curves from the Tikhonov experiments look like on 
+        datasets in which Tikhonov regularization is needed for better
+        generalization?
+    \item The part I am most interested in would be to prove high-probability
+        bounds for the wiggled estimators and potentially convergence rates
+        for the iterative algorithm (as it seems to be quite fast).
+        Luckily, I am working in a research group (MaLGa in Genova) that 
+        have very good understanding of kernel matrices of Matérn kernels
+        and have already proven many bounds for kernel ridge(less) 
+        regression in the past.
+    \item Most importantly: check the literature for any papers on these
+        methods. I haven't found anything so far but my research was not 
+        extensive.
+\end{itemize}
 
@@
 
+\bibitem{falkon}
+Rudi, Alessandro, Luigi Carratino, and Lorenzo Rosasco. "Falkon: An optimal large scale kernel method." Advances in neural information processing systems 30 (2017).
+
 \end{thebibliography}

### 2023-01-28 paper.tex
add range experiments and evaluation

@@
 \DeclareMathOperator{\Tr}{Tr}
+\DeclareMathOperator{\Te}{Te}
 \DeclareMathOperator{\argmin}{argmin}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment: Does it compare to Tikhonov?}
-
 \begin{figure}[htp]
     \centering
-    \includegraphics[width=\textwidth]{tikhonov}
-    \caption{tikhonov}
-    \label{fig:tikhonov}
+    \includegraphics[width=\textwidth]{range}
+    \caption{
+        Standard versus wiggled estimators. 
+        The green vertical line shows the base bandwidth.
+        The dots indicate the minima.
+        The blue line is indepdent of the base bandwidth. }
+    \label{fig:range}
 \end{figure}
+
+In this first experiment we want to find out the relationship between wiggled
+estimator $\hat f_{\gamma_0,\gamma}$ and standard estimator 
+$\hat f_{\gamma,\gamma}$.
+We choose a reasonably hard function $[3/2, 3] \to \R$, namely
+\[ f(x) = \zeta(x) \sin(20 x), \]
+and pick $N_{\Tr} = 30$ uniformly sampled training points.
+In a range of $\Gamma = [10,50]$ we plot the performance on 
+a training set (regular grid, $N_{\Te} = 1000$) of the "standard" estimator
+for each $\gamma$ and the "wiggled" estimator wiggled for the same 
+$\gamma$ but previously fit on a fixed base bandwidth $\gamma_0$.
+We plot this 6 times for different base bandwidth.
+See \cref{fig:range} for the results.
+
+We summarize the observations:
+\begin{itemize}
+    \item Generally, the behavior of wiggling $\gamma$ away from $\gamma_0$
+        is smooth as expected.
+    \item The first four plots show that wiggling improves over standard
+        estimators even if the base bandwidth is already optimal.
+    \item If the base bandwidth is chosen too small, then a iterated 
+        wiggle search goes in the direction of the standard optimal bandwidth.
+    \item However, plot no. 5 shows that this seems not be the case in general.
+\end{itemize}
 
@@
     \label{fig:split}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experiment: Does it compare to Tikhonov?}
+
+\begin{figure}[htp]
+    \centering
+    \includegraphics[width=\textwidth]{tikhonov}
+    \caption{tikhonov}
+    \label{fig:tikhonov}
 \end{figure}

### 2023-01-28 experiments/range.compute.py

@@ -1,0 +1,40 @@
+import numpy as np
+from sklearn.metrics import mean_squared_error
+from scipy.special import zeta
+from shared import laplace_kernel, WiggledScaleKernelLS, get_result_and_figure_path
+
+def f(x): return zeta(x) * np.sin(20*x)
+
+if __name__ == '__main__':
+    rng = np.random.default_rng(2321)
+    Ntr = 30
+    Nte = 1000
+
+    Xtr = rng.uniform(1.5, 3, Ntr).reshape(-1,1)
+    Ytr = f(Xtr)
+    Xte = np.linspace(1.5, 3, Nte).reshape(-1,1)
+    Yte = f(Xte)
+
+    base_gammas = np.array([11, 13, 17, 20, 25, 30])
+    wiggle_gammas = np.linspace(10, 50, num=100)
+    wiggled_mses = np.zeros((base_gammas.size, wiggle_gammas.size))
+
+    standard_mses = np.zeros(wiggle_gammas.size)
+    for i in range(wiggle_gammas.size):
+        model = WiggledScaleKernelLS(kernel_func=laplace_kernel, base_gamma=wiggle_gammas[i])
+        model.fit(Xtr, Ytr)
+        standard_mses[i] = mean_squared_error(Yte, model.predict(Xte))
+    min_std_mse = np.argmin(standard_mses)
+
+    for j in range(base_gammas.size):
+        model = WiggledScaleKernelLS(kernel_func=laplace_kernel, base_gamma=base_gammas[j]).fit(Xtr, Ytr)
+        for i in range(wiggle_gammas.size):
+            model.wiggled_gamma = wiggle_gammas[i]
+            Ypr = model.predict(Xte)
+            wiggled_mses[j,i] = mean_squared_error(Yte, Ypr)
+    
+    resFile,_ = get_result_and_figure_path()
+    np.savez(resFile, 
+        base_gammas=base_gammas, wiggle_gammas=wiggle_gammas, 
+        standard_mses=standard_mses, min_std_mse=min_std_mse,
+        wiggled_mses=wiggled_mses)

### 2023-01-28 experiments/range.present.py

@@ -1,0 +1,25 @@
+from shared import get_result_and_figure_path
+import numpy as np
+from matplotlib import pyplot as plt
+
+if __name__ == '__main__':
+    resFile, figFile = get_result_and_figure_path()
+    results = np.load(resFile)
+    base_gammas = results['base_gammas']
+    min_std_mse = np.argmin(results['standard_mses'])
+
+    plt.figure(figsize=(12, 6))
+    for j in range(base_gammas.size):
+        min_wig_mse = np.argmin(results['wiggled_mses'][j,:])
+        plt.subplot(2, 3, j+1)
+        plt.title('$\gamma_0 = {}$'.format(base_gammas[j]))
+        plt.xlabel('$\gamma$')
+        plt.ylabel('MSE')
+        plt.plot(results['wiggle_gammas'], results['standard_mses'], label='standard')
+        plt.scatter([results['wiggle_gammas'][min_std_mse]], [results['standard_mses'][min_std_mse]], c='blue')
+        plt.plot(results['wiggle_gammas'], results['wiggled_mses'][j,:], label='wiggled')
+        plt.scatter([results['wiggle_gammas'][min_wig_mse]], [results['wiggled_mses'][j,min_wig_mse]], c='orange')
+        plt.axvline(base_gammas[j], c='green')
+        plt.legend()
+    plt.tight_layout()
+    plt.savefig(figFile)

### 2023-01-28 experiments/shared.py

@@ -13,7 +13,7 @@ import numpy as np
 def get_result_and_figure_path():
     script = sys.argv[0]
     basename = os.path.splitext(os.path.basename(script))[0]\
-        .replace('_compute', '').replace('_present', '')
+        .replace('.compute', '').replace('.present', '')
     return ( 'build/results/' + basename + '.npz', 
         'build/figures/' + basename + '.png' )
 

### 2023-01-28 paper.tex
add and fix parts on splitting and iterating

@@
 As announced, let us fix the Laplacian kernel function
-\[ k_\gamma(x,x') = \exp\left(-\frac{\| x - x' \|}{\gamma} \right) \]
+\[ k_\gamma(x,x') = \exp\left(-\gamma \| x - x' \| \right) \]
 with bandwidth $\gamma > 0$.
-The convention here is that smaller bandwidth actually decreases the 
+The convention here is that smaller bandwidth actually increases the 
 area under the curve $x \mapsto k_0(x) \coloneqq k(0,x)$ and vice-versa.
+(different conventions everywhere!).
 This whole note is based on the observation that the kernel ridgeless 
@@
 the vectorized computation
-\[ \big( \mathbf c^{\gamma_0} \exp(-(1/\gamma) \mathbf D) \big)^T \]
+\[ \big( \mathbf c^{\gamma_0} \exp(-\gamma \mathbf D) \big)^T \]
 yielding the new predictions of the inputs in $\Tr_2$.
@@
         estimators (or if at all)?
-    \item Given that we move away from the interpolating state 
-        (in a smooth and controlled manner) does wiggling compare to 
-        Tikhonov?
-        What happens with the norm/complexitie of a wiggled estimator
-        compared to the standard one?
     \item Does an iterated wiggle search find the best bandwidth?
         Is it less expensive than classical model selection techniques?
+    %\item Given that we move away from the interpolating state 
+    %    (in a smooth and controlled manner) does wiggling compare to 
+    %    Tikhonov?
+    %    What happens with the norm/complexity of a wiggled estimator
+    %    compared to the standard one?
 \end{enumerate}
@@
     \centering
-    \includegraphics[width=\textwidth]{split}
-    \caption{split}
+    \includegraphics[width=0.8\textwidth]{split}
+    \caption{
+        Figuring out the sweet spot for the train split on real data.
+        "Baseline" always refers to the standard estimator with the 
+        base bandwidth on the full train set.
+        The test MSE curves refer to the best wiggled bandwidth. }
     \label{fig:split}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment: Does it compare to Tikhonov?}
+The last experiment gave hope that wiggle search can work well.
+In this part we will try to figure out the sweet spot between computational
+savings and performance gain.
+We move to real data now.
 
-\begin{figure}[htp]
-    \centering
-    \includegraphics[width=\textwidth]{tikhonov}
-    \caption{tikhonov}
-    \label{fig:tikhonov}
-\end{figure}
+We choose the MNIST binary classification task with digits 4 and 9
+(the hardest to distinguish) with fixed sample of size $N_{\Tr} = 1000$.
+The set $\Gamma = \Gamma(\gamma_0)$ is being chosen at each base bandwidth
+as a small logarithmic scale clustered at the center $\gamma_0$ 
+(contains 30 bandwidths) and we fix $\gamma_0 = 1/784$ which is the 
+\texttt{sklearn} default choice (based purely on dimensionality).
+The splits are percentages $p$ of $N_{\Tr}$.
+All plotted values are with respect to $p \in [0.3, 0.95]$:
+sample error, error on held-out set, error on test set, as well as the 
+chosen bandwidths and the fitting times.
+See \cref{fig:split} for the results.
+
+We summarize the observations:
+\begin{itemize}
+    \item The more data we are considering, the bigger we choose the 
+        bandwidth parameter (e.g. the smaller the area under the kernel
+        function will be) as one would expect.
+    \item We achieve equivalent performance just by using only 70\% of the
+        training set plus wiggling.
+        After that point we see only improvements in favor of wiggling.
+\end{itemize}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{iterative_algo}
+    \includegraphics[width=0.8\textwidth]{iterative_algo}
     \caption{iterative algo}
@@
 \end{figure}
+
+We have already seen in the first experiment that the direction given by
+wiggle search seems to be the right one.
+In this experiment we are going to test the iterated wiggle search as 
+described above: fit least squares, wiggle search, refit with the best bandwidth, 
+and so on.
+If the chosen bandwidth converges to something, we are likely to have hit an 
+optimum or at least something close to one.
+
+The experimental setup is the same as above.
+MNIST binary, initial bandwidth $\gamma_0 = 1/784$ and by our previous 
+findings, train split $p = 0.85$.
+We plot both the wiggled estimator's MSE with best bandwidth and that 
+of refit standard estimator with exactly that MSE.
+We also plot the history of the best bandwidth found.
+The algorithm was run for more than 9 iterations but it converged 
+and we cut the plot.
+See the results in \cref{fig:iterative_algo}.
+
+Summary:
+\begin{itemize}
+    \item The algorithm converges and decreases the MSE as hoped for.
+    \item Remarkably, the curves for wiggled and refit estimators are
+        qualitatively the same! 
+        This gives hope for finding good theoretical guarantees for this
+        method.
+    \item The wiggled estimators reliably outperform the standard ones
+        even though they are not fit on the full training set.
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\subsection{Experiment: Does it compare to Tikhonov?}
+%
 

### 2023-01-28 experiments/iterative_algo.present.py

@@ -9,7 +9,7 @@ if __name__ == '__main__':
     iters = list(range(n_iter))
 
     fig,axs = plt.subplots(1, 2)
-    fig.set_size_inches(9,4.5)
+    fig.set_size_inches(9,4)
     fig.suptitle('Average fit time: {:.2f} seconds'.format(np.mean(np.array(results['fit_times']))))
     axs[0].set_title('MSE history')
     axs[0].plot(iters, results['refit'], label='refit with best bandwidth')
@@ -19,7 +19,8 @@ if __name__ == '__main__':
     axs[0].legend()
     axs[1].set_title('bandwidth history')
     axs[1].plot(iters, results['gamma_hist'])
-    axs[0].set_xlabel('iteration')
-    axs[0].set_ylabel('$\gamma$')
+    axs[1].set_xlabel('iteration')
+    axs[1].set_ylabel('$\gamma$')
 
+    fig.tight_layout()
     fig.savefig(figFile)

### 2023-01-28 experiments/tikhonov.compute.py

@@ -1,39 +1,0 @@
-import time
-import numpy as np
-from sklearn.metrics import mean_squared_error
-from shared import MnistBinaryDataset, laplace_kernel, WiggledScaleKernelLS, WiggleSearchScaleKernelLS, get_search_gammas, get_result_and_figure_path
-
-if __name__ == '__main__':
-    Ntr = 1000
-    Nte = 1000
-    Xtr,Ytr, Xte,Yte = MnistBinaryDataset(98).sample(4, 9, [Ntr, Nte])
-
-    gamma0 = 1/784
-    alphas = np.logspace(-5, 1, num=15)
-
-    baseline = WiggledScaleKernelLS(laplace_kernel, base_gamma=gamma0)
-    baseline.fit(Xtr, Ytr)
-    baseline_mse = mean_squared_error(Yte, baseline.predict(Xte))
-    wiggled = WiggleSearchScaleKernelLS(laplace_kernel, gamma0, get_search_gammas(gamma0, 2, 20), 0.85)
-    wiggled.fit(Xtr, Ytr)
-    wiggled_mse = mean_squared_error(Yte, wiggled.predict(Xte))
-
-    tikhonov_mses = []
-    wiggled_tikhonov_mses = []
-    for i in range(alphas.size):
-        alpha = alphas[i]
-
-        startTime = time.time()
-        tikh = WiggledScaleKernelLS(laplace_kernel, base_gamma=gamma0, alpha=alpha)
-        tikh.fit(Xtr, Ytr)
-        tikhonov_mses.append(mean_squared_error(Yte, tikh.predict(Xte)))
-
-        wtikh = WiggleSearchScaleKernelLS(laplace_kernel, gamma0, get_search_gammas(gamma0, 2, 20), 0.85, alpha=alpha)
-        wtikh.fit(Xtr, Ytr)
-        wiggled_tikhonov_mses.append(mean_squared_error(Yte, wtikh.predict(Xte)))
-
-        print('finished alpha {}/{} in {:.2f} seconds'.format(i+1, alphas.size, time.time() - startTime))
-
-    resFile,_ = get_result_and_figure_path()
-    np.savez(resFile, alphas=alphas, baseline_mse=baseline_mse, wiggled_mse=wiggled_mse, 
-        tikhonov_mses=tikhonov_mses, wiggled_tikhonov_mses=wiggled_tikhonov_mses)

### 2023-01-28 experiments/tikhonov.present.py

@@ -1,18 +1,0 @@
-from shared import get_result_and_figure_path
-import numpy as np
-from matplotlib import pyplot as plt
-
-if __name__ == '__main__':
-    resFile, figFile = get_result_and_figure_path()
-    results = np.load(resFile)
-    alphas = results['alphas']
-
-    plt.plot([alphas[0], alphas[-1]], 2*[results['baseline_mse']], label='baseline')
-    plt.plot(alphas, results['tikhonov_mses'], label='tikhonov')
-    plt.plot([alphas[0], alphas[-1]], 2*[results['wiggled_mse']], label='wiggled')
-    plt.plot(alphas, results['wiggled_tikhonov_mses'], label='wiggled tikhonov')
-    plt.xscale('log')
-    plt.xlabel('$\\alpha$')
-    plt.ylabel('MSE')
-    plt.legend()
-    plt.savefig(figFile)

### 2023-01-28 paper.tex
add conclusions and strange effect appendix

@@
         Is it less expensive than classical model selection techniques?
-    %\item Given that we move away from the interpolating state 
-    %    (in a smooth and controlled manner) does wiggling compare to 
-    %    Tikhonov?
-    %    What happens with the norm/complexity of a wiggled estimator
-    %    compared to the standard one?
 \end{enumerate}
@@
     \includegraphics[width=0.8\textwidth]{iterative_algo}
-    \caption{iterative algo}
+    \caption{Iteratively applying wiggle search and then refitting
+        a standard estimator with the found optimal bandwidth.}
     \label{fig:iterative_algo}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Experiment: Does it compare to Tikhonov?}
-%
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Finally, I propose a list of todos and questions for further research on 
+This paper proposes two new methods for leveraging the bandwidth in 
+Laplacian kernel ridgeless regression.
+The results seem to indicate that there is potential for them to replace
+classical methods for optimizing the parameter (grid search, 
+cross-validation) as they are expensive and don't scale well to big data.
+The methods have been demonstrated to improve over standard estimators 
+in some cases even with nearly optimal bandwidth guess.
+However, much further experimentation is necessary and providing 
+theoretical guarantees is thinkable as the recent progresses in the field 
+have opened new doors.
+
+Finally, I add a list of todos and questions for further research on 
 the method.
 \begin{itemize}
-    \item How far is the reach of wiggling being indicative about the refit
-        performance? That is, if I fix an initial bandwidth $\gamma_0$, 
-        what is the range of $\gamma$ such that the performance of the 
-        estimator $f_{\gamma_0, \gamma}$ can predict that of the estimator
-        $f_\gamma$? In small logarithmic local ranges our experiments have
-        shown strong similaritity of the risk curves but how far can this be 
-        streched? Here, theoretical guarantees are needed to make the method
-        pratically usable.
-    \item The MNIST dataset have been shown in \cite{justinterpolate} to 
-        be predicted best by ridgeless kernel estimators.
-        How do the curves from the Tikhonov experiments look like on 
-        datasets in which Tikhonov regularization is needed for better
-        generalization?
-    \item The part I am most interested in would be to prove high-probability
-        bounds for the wiggled estimators and potentially convergence rates
-        for the iterative algorithm (as it seems to be quite fast).
-        Luckily, I am working in a research group (MaLGa in Genova) that 
-        have very good understanding of kernel matrices of Matérn kernels
-        and have already proven many bounds for kernel ridge(less) 
-        regression in the past.
-    \item Most importantly: check the literature for any papers on these
+    \item Bigger scale experiments are necessary to prove that the method
+        really scales well (my computer couldn't do it)
+        Also because the MNIST dataset have been shown in \cite{justinterpolate} 
+        to be predicted best by ridgeless Laplacian estimators.
+        What happens with more noisy datasets?
+    \item Given that we move away from the interpolating state 
+        (in a smooth and controlled manner) does wiggling compare to 
+        Tikhonov?
+        What happens with the norm/complexity of a wiggled estimator
+        compared to the standard one?
+    \item As a mathematician I really want to leave the experimental world
+        and analyze on paper providing risk bounds.
+        There is in particular one phenomenon that I found by mistake and 
+        that I couldn't label.
+        If it is a general rule, that would imply that a wiggled estimator
+        "knows" about the the standard estimator and can be used to find its 
+        global risk minimum (always w.r.t. the bandwidth).
+        See \cref{app:strange}.
+    \item Check the literature for any papers on these
         methods. I haven't found anything so far but my research was not 
-        extensive.
+        at all exhaustive.
 \end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\appendix
+\section{A strange effect} \label{app:strange}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[htp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{strange}
+    \caption{Strange: Cheap wiggled estimators seems to know the expensive 
+        standard ones perfectly well?}
+    \label{fig:strange}
+\end{figure}
+
+In one of my other experiments I used the inverse bandwidth convention.
+It is more or less the same plot as in \cref{fig:range} just with 
+$\gamma \to 1/\gamma$ so that I had to manually edit the bandwidth 
+values to obtain readable results in the new scale.
+However, in that inverse, it appeared to me that the wiggled estimator
+"knew" perfectly well where the true optimal bandwidth of the standard 
+estimator is.
+Indeed, \cref{fig:strange} shows extrema at exactlt that best point 
+even if the base bandwidth is far away.
+I coudln't confirm that this is a big coincidence but I would like to 
+investigate that more.
 

### 2023-01-28 paper.tex
update abstract

@@
     The provided experimental results indicate that \textit{wiggling}
-    the bandwidth in this way can provide generalization and potentially 
-    be a cheap alternative to the classical Tikhonov square-norm penality.
+    the bandwidth can provide generalization and improve over standard
+    estimators.
     Finally, we propose and test a new iterative method for optimizing the 
